---
title: "Basic Job Optimisation"
output:
  html_document:
    df_print: paged
  html_notebook: default
  pdf_document: default
---


**The problem**
 
Jobs - Need a number of skills to complete each which require a number a workforce hours over a number of days
 
Workforce - Will have hours which they can work each day and one or more skills which they can perform

**Acceptance Testing**

* up to 300 jobs

* how much workforce?


```{r libs, echo=FALSE,message=FALSE, warning = FALSE, results = 'hide'}

rm(list = ls())

library(lpSolveAPI)

library(tidyverse)

```


**Lp Solve Ideas**

basic ideas of framing the problem to LP solve 

* its binary
* the decision being made is do or not do the job
* bounds would be the ammount of hours available in each skill group
* goal 1 - would be min/max job priority depending on how folk see priority
* goal 2 - use all the workforce each day regardless of prioity
* how do we mix the goals together as both are valid, what if their prioirty system is super basic creating similar decisions i.e. 1/2
* how do we penalise the optimisation for using workers that are multiple skilled


**Simple Problem**

firstly coding directly into LP solve

3 jobs 

* req 7.5 hours of skill 1 and 7.5 hours of skill 2 prioirty 2
* req 7.5 hours of skill 1 and 7.5 hours of skill 2 prioirty 1
* req 7.5 hours of skill 2 prioity 2

prioirty high = good, goal is a max

if there are only 

* 7.5 hours of skill 1 
* 15 hours of skill 2 available

should pick jobs 1 and 3

```{r simple prob, echo=FALSE,message=FALSE, warning = FALSE, results = 'hide'}

#create model
#bounds, decisions
lp <- lpSolveAPI::make.lp(2, 3)

cols <- c("job1", "job2", "job3")
rows <- c("s1", "s2")
dimnames(lp) <- list(rows, cols)

set.column(lp, 1, c(7.5, 7.5), indices = c(1,2))
set.column(lp, 2, c(7.5, 7.5), indices = c(1,2))
set.column(lp, 3, c(0, 7.5), indices = c(1,2))

#the problems is binary
lpSolveAPI::set.type(lp, 1:3, "binary")

lpSolveAPI::set.constr.type(lp, c("<=", "<="))

lpSolveAPI::set.constr.value(lp, c(7.5, 15))

lpSolveAPI::set.objfn(lp, c(2,1,2))
  
lpSolveAPI::lp.control(lp, sense = "max")

lp

model <- solve(lp)

model

lpSolveAPI::get.variables(lp)


```

done, easy!

```{r simple ouptut, echo=FALSE, message=FALSE, warning = FALSE}

lp

model <- solve(lp)

model

lpSolveAPI::get.variables(lp)


```

**What if there is a multi skill worker?**

same 3 jobs:

* req 7.5 hours of skill 1 and 7.5 hours of skill 2 prioirty 2
* req 7.5 hours of skill 1 and 7.5 hours of skill 2 prioirty 1
* req 7.5 hours of skill 2 prioity 2

workers are now:

* worker 1 has 7.5 hours of skill 1 or 2
* workers 2 and 3 can only do skill 2

should still be same answer but need to account for that worker 1 can do all 3 jobs

first just the first 2 jobs so can see the model on screen


```{r multi skill, echo=FALSE,message=FALSE, warning = FALSE, results = 'hide'}

#create model
lp <- lpSolveAPI::make.lp(nrow = 5, ncol = 6)


cols <- c("j1", "j1_s1_s1", "j1_s2_s12", "j1_s2_s2", 
          "j2", "j2_s1_s1", "j2_s2_s12", "j2_s2_s2")
rows <- c("s12", "s2", 
          "job1_s2", "job1", 
          "job2_s2", "job2")        
          
lp <- lpSolveAPI::make.lp(nrow = length(rows), ncol = length(cols))

dimnames(lp) <- list(rows, cols)


#first decision is NOT doing the job
set.column(lp, 1, c(2), indices = c(4))


#first decision is job 1 skill 1, only 1 person can do this
set.column(lp, 2, c(7.5, 1), indices = c(1,4))


#next 2 decisions are job 1 skill 1/2 and skill 2
#technically if there was only 1 person to do skill 1 2 decisions would not be needed
set.column(lp, 3, c(7.5, 0, 1, 1), indices = c(1,2,3,4))
set.column(lp, 4, c(0, 7.5, 1, 1), indices = c(1,2,3,4))


#job 2 is a repeat of job 1

set.column(lp, 5 , c(2), indices = c(6))

set.column(lp, 6, c(7.5, 1), indices = c(1,6))

set.column(lp, 7, c(7.5, 0, 1, 1), indices = c(1,2,5,6))
set.column(lp, 8, c(0, 7.5, 1, 1), indices = c(1,2,5,6))


#the problems is binary
lpSolveAPI::set.type(lp, 1:length(cols), "binary")




lpSolveAPI::set.constr.type(lp, 
                            types = c("<=", "<=", 
                                      "<=", "=", 
                                      "<=", "="))
                                      
                                     # "<="))#, 
                                      #"<=", "="))


lpSolveAPI::set.constr.value(lp, rhs = c(7.5, 15, 
                                         1, 2, 
                                         1, 2)) # , 
                                         #1, 2))



lpSolveAPI::set.objfn(lp, c(0,1,1,1,
                            0,0.5,0.5,0.5))
                            
                          #  ,
                          #  0,0.2,0.2,0.2))

lpSolveAPI::lp.control(lp, sense = "max")

model <- solve(lp)

model

lpSolveAPI::get.variables(lp)

lp


```

The structure is

* each decision represents matching the required skill of the job to the skill groups available from the workforce
* if there is only 1 match e.g. you need Skill 1 and there is only 1 group of people who can do Skill 1 then this is 1 decision as seen below with j1_s1_s1 and j2_s1_s1, this is either on or off
* where there are 2 or more groups of people that can satisfy the required skill each one is represented as a decision and they are bounded so that only one can be selected e.g. j1_s2_s12 and j1_s2_s2 bounded by job1_s2 being <= 1 
* where multiple skills are required for a job, a further decision is added for NOT doing the job and a bound is added that it has to pick not doing or doing the job. below bound job1 ensures that either the job is not done (with a goal of 0) it is done which is a combination of j1_s1_s1 + j1_s2_s12 or j1_s1_s1 + j1_s2_s2. This bound is set as = to the total number of skills required to do the job which in the below example is 2.


```{r multi skill output 1, echo=FALSE, message=FALSE, warning = FALSE}

lp

```

does pick the correct decisions, but the output is quite confusing, at moment you need to match the 1's and 0's to the above columns

```{r multi skill output 2, echo=FALSE, message=FALSE, warning = FALSE}

model <- solve(lp)

model

lpSolveAPI::get.variables(lp)


```

**Creating a basic function to create the model from input data**

```{r create basic functions, echo=FALSE,message=FALSE, warning = FALSE, results = 'hide'}

jobs <- tibble(J_ID = c(1,2), Priority = c(2,1) )

jobs_skills <- tibble(J_ID = c(1,1,2,2), Skill = c("S1","S2","S1","S2"), Hours = c(7.5,7.5,7.5,7.5) )

worker_skills <- tibble(E_ID = c(1,1,2,3), Skill = c("S1","S2","S2","S2"))

worker_hours <- tibble(E_ID = c(1,2,3), Hours = c(7.5, 7.5,7.5))


create_model_data <- function(jobs, jobs_skills, worker_skills, worker_hours){
  
  #first need to create how much of each type of skill is available

  ###### should be done on data load in app????? ---------------------------

  # number each available skill
  skills_id <- worker_skills %>%
    distinct(Skill) %>%
    rowid_to_column("S_ID")

  # apply numbering to worker skills table
  worker_skills <- left_join(worker_skills, skills_id, by = "Skill") %>%
    group_by(E_ID) %>%
    mutate(S_ID_G = paste0(S_ID, collapse = "")) %>%
    mutate(S_ID_G = paste0("S", S_ID_G)) %>%
    ungroup()
  
  ################ function would start here
  
  #get bounds on worker skills
  model_bounds_worker <- worker_skills %>%
    left_join(worker_hours, by = "E_ID") %>%
    distinct(E_ID, S_ID_G, Hours) %>%
    group_by(S_ID_G) %>%
    summarise(Value = sum(Hours)) %>%
    mutate(B_ID = row_number())
  
  model_bounds_value <- model_bounds_worker$Value
  model_bounds_type <- rep("<=", nrow(model_bounds_worker))
  
  
  #get all the possible worker skill combos
  worker_skills_combo <- distinct(worker_skills, S_ID, S_ID_G)
  
  #apply skills and worker skills combos to all jobs
  model_columns <- jobs_skills %>%
    left_join(skills_id, by = "Skill") %>%
    left_join(worker_skills_combo, by = "S_ID") %>%
    rename(Value = Hours)
  
  #add bound id for skill to each model column
  model_columns <- model_columns %>%
    left_join( select(model_bounds_worker, S_ID_G, B_ID), by = "S_ID_G" )
  
  
  # does the job require a do not do decision as there are multiple skills required ------------------------
  # also create a B_ID_J so that the bound can be placed in the correct position
  multiple_options_all <- model_columns %>%
    group_by(J_ID, S_ID) %>%
    summarise(count = n()) %>%
    ungroup()
  
  multiple_options <- multiple_options_all %>%
     group_by(J_ID) %>%
     summarise(max = n()) %>%  #the total number which needs to be picked
     filter(max >= 2) %>%
     rowid_to_column("B_ID_J") %>%
     mutate(
        B_ID_J = B_ID_J + max(model_bounds_worker$B_ID)
     )
    
  # create a column for each do not pick option
  model_columns_dn <- group_by(model_columns, J_ID) %>% filter(row_number()==1) %>%
    right_join(multiple_options, by = "J_ID") %>%
    mutate(Value = max,
           S_ID = 0,
           Skill = "",
           S_ID_G = 0,
           B_ID = 0) %>%
    select(-max, - B_ID_J) %>%
    ungroup()
    
  # add back into main column data
  model_columns <- rbind(model_columns, model_columns_dn) %>%
    arrange(J_ID, S_ID_G)
  
  
  # extra bounds required to act of the whole job being selected
  model_columns <- model_columns %>%
    left_join(select(multiple_options, J_ID, B_ID_J), by = "J_ID" )
  
  #update bound strings
  model_bounds_value <- c(model_bounds_value, multiple_options$max)
  model_bounds_type <- c(model_bounds_type, rep("=", nrow(multiple_options)))
  
  
  #do any bounds need adding so that only 1 of the options based on skills can be selected? --------------
  multiple_options <- multiple_options_all %>%
    filter(count >=2) %>%
    select(-count) %>%
    rowid_to_column("B_ID_JS") %>%
    mutate(
      B_ID_JS = B_ID_JS + max(model_columns$B_ID_J, na.rm = TRUE)
    )
  
  
  model_columns <- model_columns %>%
    left_join(multiple_options, by = c("J_ID", "S_ID")) %>%
    replace(is.na(.), 0)
  
  model_bounds_type_temp <- rep("<=", nrow(multiple_options))
  
  model_bounds_value <- c(model_bounds_value, rep(1, length(model_bounds_type_temp)))
  model_bounds_type <- c(model_bounds_type, model_bounds_type_temp)
  
  
  #create final id for the columns
  model_columns <- model_columns %>%
    rowid_to_column("C_ID")
  
  
  # add goal as a column
  model_columns <- model_columns %>%
    left_join(jobs, by = "J_ID" )
  
  # if there is a do not pick position change priority to 0
  model_columns <- model_columns %>%
    mutate(
      Priority = ifelse(S_ID == 0, 0 , Priority)
    )

  
  return( 
    list(
      model_columns = model_columns,
      model_bounds_value = model_bounds_value,
      model_bounds_type = model_bounds_type
    )
  )
  
  
}


run_model <- function(model_data_list, model_return = FALSE){

  # create model 
  model_columns <- model_data_list$model_columns
  
  columns <- nrow(model_columns)

  rows <- max(model_columns$B_ID_JS)

  lp <- lpSolveAPI::make.lp(nrow = rows, ncol = nrow(model_columns))
  
  for(i in 1:columns){
  
    col <- model_columns[i,]
  
    # get all positions
    position <- c(col$B_ID, col$B_ID_J, col$B_ID_JS)
  
    # remove zeros
    position <- position[position!=0]
  
    # create a vector of 1's based on the numbee of positons
    # all positions apart from the first are just 1's
    values <- rep( 1, length(position) )
  
    # replace the first value with the value from the column
    # if this is the do not pick position it will be the number that has to be picked
    #if it is a work option it will be number of hours
    values[1] <- col$Value
  
    set.column(lp, col$C_ID, values, indices = position)
  
  }
  
  #the problems is binary
  lpSolveAPI::set.type(lp, 1:columns, "binary")

  lpSolveAPI::set.constr.type(lp, 
                            types = model_data_list$model_bounds_type)
                                      

  lpSolveAPI::set.constr.value(lp, rhs = model_data_list$model_bounds_value)


  lpSolveAPI::set.objfn(lp, model_columns$Priority)


  lpSolveAPI::lp.control(lp, sense = "max")


  model <- solve(lp)
  
  if (!model_return) {
    lp <- ""
  }
  
  results <-  lpSolveAPI::get.variables(lp) 
  
  model_columns <- cbind(model_columns, results)

  return(
    list(
      model_columns = model_columns,
      lp = lp
    )
  )
    
}


```

data looks like this

jobs have 2 files

first the jobs themselves, unique id and a priority

```{r jobs, echo=FALSE, message=FALSE, warning = FALSE}

jobs

```

second the skills and hours that are required for each job

```{r jobs skills, echo=FALSE, message=FALSE, warning = FALSE}

jobs_skills

```

for the workers there are also 2 files

first for each worker how many hours do they have

```{r worker_hours, echo=FALSE, message=FALSE, warning = FALSE}

worker_hours

```

second, what skills does each worker have

```{r worker_skills, echo=FALSE, message=FALSE, warning = FALSE}

worker_skills

```

model looks the same

everything has been added programatically so there are some differences

the names of the columns and rows have been added to help understanding

* decisions are still in the same order left to right
* rows (bounds) are in the order skills, jobs, skills within jobs


```{r function output, echo=FALSE, message=FALSE, warning = FALSE}

model_data_list <- create_model_data(jobs, jobs_skills, worker_skills, worker_hours)

results <- run_model(model_data_list, TRUE)

lp <- results$lp


cols <- c("j1", "j1_s1_s1", "j1_s2_s12", "j1_s2_s2", 
          "j2", "j2_s1_s1", "j2_s2_s12", "j2_s2_s2")
rows <- c("s12", "s2", 
          "job1", "job2",
          "job1_s2", "job2_s2" )        
          
dimnames(lp) <- list(rows, cols)

lp

```

and so do the results in that it picks job 1

```{r multi skill output results, echo=FALSE, message=FALSE, warning = FALSE}

results$model_columns$results

```

using a super quick basic graphing function

```{r basic graph, echo=FALSE,message=FALSE, warning = FALSE, results = 'hide'}


quick_graph <- function(worker_hours, results_table){

  #super quick (shit) function to visualise results
  results_table <- results_table %>%
    filter(results == 1, Skill != "") %>%
    group_by(group) %>%
    summarise(Value = sum(Value)) %>%
    mutate(day = "Today")
  
  worker_hours_table <- worker_hours %>%
    mutate(day = "Today") %>%
    group_by(day) %>%
    summarise(Value = sum(Hours))
  
  plot <- ggplot(results_table, aes(day, Value))
  
  plot <- plot + geom_bar(aes(fill=group), width = 0.5, stat="identity") + 
    labs(title="Summary") +
    geom_point(aes(x=worker_hours_table$day, y=worker_hours_table$Value), stat="identity", size = 4)

  return(plot)
    
}

quick_graph(worker_hours, mutate(results$model_columns, group = Skill))

quick_graph(worker_hours, mutate(results$model_columns, group = paste0("Job", J_ID)))


```

same problem as before but with 3 jobs, should pick jobs 1 and 3

```{r}


jobs <- tibble(J_ID = c(1,2,3), Priority = c(2,1,2) )
 
jobs_skills <- tibble(J_ID = c(1,1,2,2,3), Skill = c("S1","S2","S1","S2","S2"), Hours = c(7.5,7.5,7.5,7.5,7.5) )

model_data_list <- create_model_data(jobs, jobs_skills, worker_skills, worker_hours)

results <- run_model(model_data_list, TRUE)

#quick_graph(worker_hours, mutate(results$model_columns, group = Skill))

quick_graph(worker_hours, mutate(results$model_columns, group = paste0("Job", J_ID)))


```

add in someone who can only do skill 1 and someone who can do S2/S3. should mean all 3 jobs can now be done

```{r}


worker_hours <- tibble(E_ID = c(1,2,3,4,5), Hours = c(7.5, 7.5, 7.5, 7.5, 7.5))

worker_skills <- tibble(E_ID = c(1,1,2,3,4,5,5), Skill = c("S1","S2","S2","S2","S1","S2","S3"))

model_data_list <- create_model_data(jobs, jobs_skills, worker_skills, worker_hours)

results <- run_model(model_data_list, TRUE)

#quick_graph(worker_hours, mutate(results$model_columns, group = Skill))

quick_graph(worker_hours, mutate(results$model_columns, group = paste0("Job", J_ID)))


```

add in another job with high priority that requires S3

job 2 is serviced by someone who can do S1 and S2/S3 

so a higher priority job that requires S3 means that job 2 drops out

maybe need to think about configuring the bounds to favour a mix of high priority/filling all jobs

```{r}

jobs <- tibble(J_ID = c(1,2,3,4), Priority = c(2,1,2,2) )
 
jobs_skills <- tibble(J_ID = c(1,1,2,2,3,4), Skill = c("S1","S2","S1","S2","S2","S3"), Hours = c(7.5,7.5,7.5,7.5,7.5,7.5) )

model_data_list <- create_model_data(jobs, jobs_skills, worker_skills, worker_hours)

results <- run_model(model_data_list, TRUE)

#quick_graph(worker_hours, mutate(results$model_columns, group = Skill))

quick_graph(worker_hours, mutate(results$model_columns, group = paste0("Job", J_ID)))


```

test with 3 skills required for one job, job 1 now requires skill 3 so its a higher priority than the previously added job 4 above so now should pick jobs 1 and 3.

```{r}

jobs <- tibble(J_ID = c(1,2,3,4), Priority = c(2,1,2,2) )
 
jobs_skills <- tibble(
  J_ID = c(1,1,2,2,3,4,1), 
  Skill = c("S1","S2","S1","S2","S2","S3","S3"), 
  Hours = c(7.5,7.5,7.5,7.5,7.5,7.5,7.5) 
  )

model_data_list <- create_model_data(jobs, jobs_skills, worker_skills, worker_hours)

results <- run_model(model_data_list, TRUE)

#quick_graph(worker_hours, mutate(results$model_columns, group = Skill))

quick_graph(worker_hours, mutate(results$model_columns, group = paste0("Job", J_ID)))


```

add job 5 thats of a lower priority but requires skill 1 so will plug the gap and now pick jobs 1,3 and 5

```{r}

jobs <- tibble(J_ID = c(1,2,3,4,5), Priority = c(2,1,2,2,1) )
 
jobs_skills <- tibble(
  J_ID = c(1,1,2,2,3,4,1,5), 
  Skill = c("S1","S2","S1","S2","S2","S3","S3","S1"), 
  Hours = c(7.5,7.5,7.5,7.5,7.5,7.5,7.5,7.5) 
  )

model_data_list <- create_model_data(jobs, jobs_skills, worker_skills, worker_hours)

results <- run_model(model_data_list, TRUE)

#quick_graph(worker_hours, mutate(results$model_columns, group = Skill))

quick_graph(worker_hours, mutate(results$model_columns, group = paste0("Job", J_ID)))


```

